Is high social class always beneficial for survival? Northern Sweden 1801--2013

### 2018-04-04 marginal.Rmd
fix random error in Model

The comparison of mortality in the five social classes, period by period, is done
by comparing estimated survival curves, adjusted for the distribution of the confounding variables over the social classes.
The adjustment carried out by inverse probability weighting (**reference**) in the following way: First, for each social class, 
the probability that an individual in the class belongs to that class is estimated by logistic regression with the 
confounders as explanatory variables. Second, these probabilities are inverted and used as weights in the estimation of the survival 
curves for each social class.

@@

-->

The disturbance effect of confounders is 
balanced out through weighting. In our case , `civil status` and `urban/rural` are confounders, that is, they 
are correlated to both our outcome, mortality, and our explanatory variable, social class. The latter association 
manifests itself through the facts that the marriage frequency varies by social class and certain social classes are 
more common in the rural area than others (we find farmers on the countryside).

@@

plot(c(1812, 1838, 1862, 1888, 1912, 1938, 1978, 1982, 1988, 1995, 2005, 2010), elite / kk, type = "b", col = "black", lty = 1, pch = 1, ylim = c(0, 4))
lines(c(1812, 1838, 1862, 1888, 1912, 1938, 1978, 1982, 1988, 1995, 2005, 2010), middle / kk, type = "b", col = "green", lty = 2, pch = 2)
lines(c(1812, 1838, 1862, 1888, 1912, 1938, 1978, 1982, 1988, 1995, 2005, 2010), worker / kk, type = "b", col = "blue", lty = 3, pch = 3)
lines(c(1812, 1838, 1862, 1888, 1912, 1938, 1978, 1982, 1988, 1995, 2005, 2010), farmer / kk, type = "b", col = "darkgreen", lty = 4, pch = 4)
lines(c(1812, 1838, 1862, 1888, 1912, 1938, 1978, 1982, 1988, 1995, 2005, 2010), no / kk, type = "b", col = "red", lty = 5, pch = 5)
```

>>>>

The comparison of mortality in the five social classes, period by period, is done
by comparing estimated survival curves, adjusted for the distribution of the confounding variables over the social classes.
The adjustment carried out by inverse probability weighting [@hoth52;@hero06] in the following way: First, for each social class, 
the probability that an individual in the class belongs to that class is estimated by logistic regression with the 
confounders as explanatory variables. Second, these probabilities are inverted and used as weights in the estimation of the survival 
curves for each social class.

The social classes are compared by the \emph{expected years of life
before 65(100)} for a person aged 40(65). This measure is calculated
as the area under the survival curve for each class.

@@

-->
<!--
The disturbance effect of confounders is 
balanced out through weighting. In our case , `civil status` and `urban/rural` are confounders, that is, they 
are correlated to both our outcome, mortality, and our explanatory variable, social class. The latter association 
manifests itself through the facts that the marriage frequency varies by social class and certain social classes are 
more common in the rural area than others (we find farmers on the countryside).

@@

plot(c(1812, 1838, 1862, 1888, 1912, 1938, 1978, 1982, 1988, 1995, 2005, 2010), elite / kk, type = "b", col = "black", lty = 1, pch = 1, ylim = c(0, 4))
lines(c(1812, 1838, 1862, 1888, 1912, 1938, 1978, 1982, 1988, 1995, 2005, 2010), middle / kk, type = "b", col = "green", lty = 2, pch = 2)
lines(c(1812, 1838, 1862, 1888, 1912, 1938, 1978, 1982, 1988, 1995, 2005, 2010), worker / kk, type = "b", col = "blue", lty = 3, pch = 3)
lines(c(1812, 1838, 1862, 1888, 1912, 1938, 1978, 1982, 1988, 1995, 2005, 2010), farmer / kk, type = "b", col = "darkgreen", lty = 4, pch = 4)
lines(c(1812, 1838, 1862, 1888, 1912, 1938, 1978, 1982, 1988, 1995, 2005, 2010), no / kk, type = "b", col = "red", lty = 5, pch = 5)
```
-->
